---
title: 'Part3: carbon cost'
author: "Lei Song"
date: "1/4/2023"
toc: true
toc-depth: 2
number-sections: false
toc-location: left
output: 
  html:
      theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

# Load libraries
library(sf)
library(here)
library(terra)
library(dplyr)
library(purrr)
select <- dplyr::select

# Set paths
carbon_data_dir <- here("data/carbon")
trdeoff_dir <- here("data/tradeoff")
```

## Introduction

In this study, we used a synthesis of multiple datasets to estimate carbon cost for land conversion. The overall carbon includes above- and below-ground carbon. The used datasets are:

-   [SOC stocks (0-30 cm) made by FAO](http://54.229.242.119/GSOCmap/)
-   [SoilGrids 2.0](https://files.isric.org/soilgrids/latest/)
-   [NCEO Africa Aboveground Biomass map for 2017](https://leicester.figshare.com/articles/dataset/Africa_Aboveground_Biomass_map_for_2017/15060270)
-   [CCI BIOMASS 2018](https://data.ceda.ac.uk/neodc/esacci/biomass/data/agb/maps/v3.0/geotiff/2018)

The carbon cost is calculated with two parts:

**Soil organic carbon stock in the top 1 m**, which is usually calculated by multiplying a factor with soil organic carbon stock. The reference equation is as follows:

$$SOC\,stock = H * BD * OC $$ where $SOC\,stock$ is the soil organic carbon stock, $H$ is soil depth, $BD$ is the bulk density, and $OC$ is the soil organic carbon concentration.

**Above- and below-ground biomass**, which is the sum of above-ground biomass and an estimated below-ground biomass by multiplying a root-to-shoot ratio with the above-ground biomass.

There are a few root-to-shoot ratios used in various studies that are relevant to Tanzania.

-   0.42, Miombo woodland of Mozambique, Ryan et al. 2011.
-   0.642, Savanna; 1.887, tropical/subtropical grassland, Mokany et al. 2006.
-   0.205/0.235 (<125 Mg/ha), tropical/subtropical moist forest, Mokany et al. 2006.

The ratio of converting biomass to carbon:

-   0.5, Saatchi et al. 2011
-   0.49, Liang et al. 2021

## Soil carbon stock

We calculated soil carbon stock by combining GSOCmap v1.6.1 and SoilGrids 2.0.

```{r}
bry <- read_sf(here("data/geoms/mainland_tanzania.geojson")) %>% select()

# Base URL for ISRIC SoilGrid250m
sg_url <- "/vsicurl?max_retry=3&retry_delay=1&list_dir=no&url=https://files.isric.org/soilgrids/latest/data/"

# Set the bounding box for download
sf_use_s2(FALSE)
crs_igh <- rast(paste0(sg_url,'soc/soc_0-5cm_mean.vrt')) %>% crs()
crs_wgs <- "EPSG:4326"
bbox_igh <- bry %>% st_transform(crs_igh) %>% st_bbox() %>% 
    st_as_sfc() %>% st_buffer(1e5)
bbox_ll <- bry %>% st_bbox() %>% 
    st_as_sfc() %>% st_buffer(1)

# Download
depths <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
items <- c("soc", "bdod")
sg_dir <- file.path(carbon_data_dir, "SoilGrid250m")
if (!dir.exists(sg_dir)) dir.create(sg_dir)

# Download SOC and BDOD
walk(items, function(item){
    dst_dir <- file.path(sg_dir, item)
    if (!dir.exists(dst_dir)) dir.create(dst_dir)
    
    # For each depth
    walk(depths, function(depth){
        nm <- sprintf('%s/%s_%scm_mean.vrt', item, item, depth)
        lyr <- rast(
            paste0(sg_url, nm)) %>% 
            crop(bbox_igh) %>% project(crs_wgs)
        writeRaster(lyr, file.path(dst_dir, basename(gsub("vrt", "tif", nm))))
    })
})

# Calculate the soil organic carbon stock at 1 m
# we used the OCS at 0-30cm made by SoilGrid + GSOCmap
# plus the calculated OCS at 30-100cm.
# SoilGrid, bulk density: cg/cm3
# SoilGrid, SOC, dg/kg
# organic carbon
fnames <- file.path(sg_dir, sprintf("soc/soc_%scm_mean.tif", depths))
socr <- lapply(fnames, function(fname) {
  r <- rast(fname)
  # convert dg/kg to g/kg to percent
  r / 10 / 10
})

# bulk density
fnames <- file.path(sg_dir, sprintf("bdod/bdod_%scm_mean.tif", depths))
bd <- lapply(fnames, function(fname) {
  r <- rast(fname)
  # convert from cg/cm3 to kg/dm3 (or g/cm3)
  r  / 100
})

# layer depth
depths <- c(0, 5, 15, 30, 60, 100, 200)
thickness <- depths[-1] - depths[-length(depths)]

# calculate carbon in each layer
soct <- do.call(c, (lapply(1:length(thickness), function(x) {
  socr[[x]] * bd[[x]] * thickness[x]
})))
names(soct) <- gsub("_mean", "", gsub("soc", "ocs", names(soct)))
writeRaster(soct, file.path(carbon_data_dir, "ocs_depths_sg.tif"))

# calculate soil organic carbon stock at 1m depth
## 0-30cm, use the mean of OCS directly made by SoilGrid250m
## and the GSOCmap 1.6.1
nm <- sprintf('%s/%s_%scm_mean.vrt', "ocs", "ocs", "0-30")
ocs_30_sg <- rast(
    paste0(sg_url, nm)) %>% 
    crop(bbox_igh) %>% project(crs_wgs)
writeRaster(ocs_30_sg, file.path(carbon_data_dir, "ocs_30cm_sg.tif"))

gsoc <- rast(file.path(carbon_data_dir, "GSOCmap1.6.1.tif")) %>% 
    crop(bbox_ll) %>% 
    resample(ocs_30_sg)
ocs_30 <- (ocs_30_sg + gsoc) / 2

## 30-100cm, calculate with soil organic carbon content and bulk density
ocs_30_100 <- sum(subset(soct, 4:5))

## 1m, sum them up
ocs_1m <- ocs_30 + ocs_30_100
writeRaster(soct_1m, file.path(carbon_data_dir, "ocs_1m_sg.tif"))
```

## Vegetation carbon stock

We determined the amount of carbon stored in the vegetation above the ground by the weighted average of the National Carbon Emissions Operations (NCEO) and the Climate Change Initiative (CCI) BIOMASS. We then multiplied above-ground biomass and different land cover root-to-shoot ratios to calculate the amount of carbon stored in the roots and other below-ground parts of the vegetation.

```{r}
# CCI BIOMASS, Mg/ha
fnames <- list.files(file.path(carbon_data_dir, "CCI_BIOMASS"),
                     full.names = TRUE)
sd_index <- grepl("SD", fnames)
abv_bms_cci <- do.call(
    merge, lapply(fnames[!sd_index], function(fname){
    rast(fname)
})) %>% crop(bbox_ll)
abv_bms_sd_cci <- do.call(
    merge, lapply(fnames[sd_index], function(fname){
    rast(fname)
})) %>% crop(bbox_ll)

# NCEO, Mg/ha
fnames <- list.files(file.path(carbon_data_dir, "NCEO"),
                     full.names = TRUE)
sd_index <- grepl("SD", fnames)
abv_bms_nceo <- rast(fnames[!sd_index]) %>% crop(bbox_ll) %>% 
    resample(abv_bms_cci)
abv_bms_sd_nceo <- rast(fnames[sd_index]) %>% crop(bbox_ll) %>% 
    resample(abv_bms_cci)

# Weighted average
factor_cci <- (abv_bms_cci + 0.0001) / (abv_bms_sd_cci + 0.001)
factor_nceo <- (abv_bms_nceo + 0.0001) / (abv_bms_sd_nceo + 0.001)
abv_bms <- (abv_bms_cci * factor_cci + abv_bms_nceo * factor_nceo) / 
    (factor_cci + factor_nceo)
writeRaster(abv_bms, file.path(carbon_data_dir, "aboveground_biomass.tif"))

# Below-ground biomass
# 0.42, Miombo woodland of Mozambique, Ryan et al. 2011.
# 0.642, Savanna; 1.887, tropical/subtropical grassland, Mokany et al. 2006.
# 0.205/0.235 (<125 Mg/ha), tropical/subtropical moist forest, Mokany et al. 2006.
# 0.563/0.275 (<20 Mg/ha), tropical/subtropical dry forest, Mokany et al. 2006.
# so, savanna - (0.42 + 0.642) / 2
# tropical dry forest - < 20, 0.563, > 20, 0.275
# grassland (plus wetland and bareland) - 1.887
# cropland - 0.2 (estimated)
# water/built-up - 0
landcover <- rast(file.path("data/landcover/landcover_tz_1km.tif"))
landcover <- project(landcover, abv_bms, method = "near")

# savanna
msk <- landcover == 4
blw_bms_shrub <- abv_bms * msk * ((0.42 + 0.642) / 2)

# tropical dry forest
msk <- landcover == 2 & abv_bms <= 20
blw_bms_tree1 <- abv_bms * msk * 0.563
msk <- landcover == 2 & abv_bms > 20
blw_bms_tree <- abv_bms * msk * 0.275 + blw_bms_tree1

# grassland
msk <- landcover %in% c(3, 8, 7)
blw_bms_grass <- abv_bms * msk * 1.887

# cropland, calculate it as well to keep the map continuous
msk <- landcover == 1
blw_bms_crop <- abv_bms * msk * 0.2

# water/built-up
msk <- landcover %in% c(5, 6)
blw_bms_wb <- abv_bms * msk * 0

blw_bms <- blw_bms_shrub + blw_bms_tree + 
    blw_bms_grass + blw_bms_crop + blw_bms_wb
writeRaster(blw_bms, file.path(carbon_data_dir, "belowground_biomass.tif"))

# Both above and below-ground biomass
root_shoot <- 0.49
bms_carbon <- (abv_bms + blw_bms) * root_shoot
writeRaster(bms_carbon, file.path(carbon_data_dir, "biomass_carbon.tif"))
```

## All carbon stock

```{r}
# soil carbon: t/ha
# biomass carbon: Mg/ha
# they are the same
## 100% of vegetative carbon and 25% of soil carbon
carbon <- bms_carbon + resample(soct_1m, bms_carbon) * 0.25
names(carbon) <- "carbon_density"
writeRaster(carbon, file.path(carbon_data_dir, "carbon_density.tif"))
```

## Carbon cost

```{r}
# Carbon cost = carbon density * convert area
# t/ha * ha = t
plant_area <- rast("data/landcover/plant_area_tz.tif")
# Each pixel is 1km * 1km = 100ha,
plant_area <- plant_area * 100
plant_area[is.na(plant_area)] <- 0

# Resample
carbon <- resample(carbon, plant_area)

area_release_80 <- 100 * 0.8 - plant_area
area_release_80[area_release_80 < 0] <- 0
carbon_cost_80 <- carbon * area_release_80
writeRaster(carbon_cost_80, file.path(trdeoff_dir, "carbon_cost_80.tif"))

area_release_60 <- 100 * 0.6 - plant_area
area_release_60[area_release_60 < 0] <- 0
carbon_cost_60 <- carbon * area_release_60
writeRaster(carbon_cost_60, file.path(trdeoff_dir, "carbon_cost_60.tif"))
```

