---
title: "Figures for manuscript"
author: "Lei Song"
format: html
---

```{r}
library(terra)
library(sf)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(tidyterra)
library(stringr)
library(dplyr)
library(colorspace)
library(stars)
library(rnaturalearth)
library(ggfx)
library(stringr)
sf_use_s2(FALSE)

# Paths
agro_dir <- "data/agriculture"
lc_dir <- "data/landcover"
tdf_dir <- "data/tradeoff"

# Study area
bry <- read_sf("data/geoms/mainland_tanzania.geojson") %>% 
    select() %>% mutate(name = "Tanzania")
```

## Study area

### Figure 1

```{r}
# Habitats
## Read palette
color_table <- readLines(
    file.path(lc_dir, "lc_legend_8classes.qml"))
color_table <- color_table[str_detect(color_table, "paletteEntry")]
color_table <- do.call(rbind, lapply(color_table, function(line){
    line <- gsub("<paletteEntry ", "", line)
    items <- strsplit(line, '\" ')[[1]]
    label <- items[str_detect(items, "label")]
    label <- gsub('        label=\"','', label)
    color <- items[str_detect(items, "color")]
    color <- gsub('[^[:alnum:] ]|color','', color)
    color <- paste0("#", color)
    value <- items[str_detect(items, "value")]
    value <- as.numeric(gsub('[^[:alnum:] ]|value','', value))
    data.frame(value = value, color = color, label = label)}))
color_table$label[2] <- "Forest/dense tree"

landcover <- rast(file.path(lc_dir, "landcover_tz_1km.tif")) %>% 
    project("epsg:4326", method = "near") %>% 
    crop(bry) %>% mask(bry)
names(landcover) <- "landcover"

# Convert map to factors
cls <- data.frame(id = color_table$value, 
                  landcover = color_table$label)
levels(landcover) <- cls

coltab(landcover) <- data.frame(
    value = color_table$value, col = color_table$color)
landcover <- st_as_stars(landcover)

# Protected areas
pas <- read_sf(
    file.path("data/protected_area", "WDPA_WDOECM_Jan2023_Public_TZA",
              "WDPA_WDOECM_Jan2023_Public_TZA.gdb"),
    layer = "WDPA_WDOECM_poly_Jan2023_TZA")
pas <- pas %>% 
    select(WDPAID, NAME, DESIG) %>% 
    rename(Geometry = SHAPE) %>% 
    filter(!DESIG %in% c("Marine Reserve", "Marine Park"))
pas <- pas %>% 
    slice(unique(unlist(st_intersects(bry, pas)))) %>% 
    st_simplify()

# Make the figure
## Overview
countries <- ne_countries(
    type = 'countries', scale = 'large',
    continent = "Africa", returnclass = 'sf')
east_africa <- read_sf(file.path("data/geoms/east_africa.sqlite"))

# Make overview
overview <- ggplotGrob(ggplot() +
    with_shadow(
        geom_sf(data = countries, fill = 'lightgrey', 
                color = 'black', linewidth = 0.3),
        sigma = 1, x_offset = 10, y_offset = 10) +
    geom_sf(data = bry, fill = 'orange', 
            color = 'grey', linewidth = 0.3) +
    coord_sf(xlim = c(-17.892, 51.892), 
             ylim = c(-35.321, 37.715)) +
    theme_transparent() +
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major = element_blank(),
          plot.margin = unit(rep(-0.6, 4), "cm")))

## Main figure
ggplot() +
    geom_stars(data = landcover, na.action = na.omit, alpha = 0.9) +
    scale_fill_manual(
        "Major land cover types", values = color_table$color,
        guide = guide_legend(
            override.aes = list(linetype = "blank", 
                                shape = NA),
            nrow = 3, title.position = "top", 
            order = 1, title.hjust = 0.5)) +
    geom_sf(data = pas, aes(color = "Protected areas"), 
            fill = "transparent", linewidth = 0.4) +
    geom_sf(data = bry, aes(color = "Mainland Tanzania"), 
            fill = "transparent", linewidth = 0.6) +
    scale_color_manual(
        NULL, values = c("grey", "black"),
        breaks = c("Mainland Tanzania", 'Protected areas'),
        guide = guide_legend(order = 2)) +
    annotation_custom(grob = overview, 
                      xmin = 37.2, xmax = 40.1,
                      ymin = - 3.8, ymax = -1.4) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10, face = "bold"),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.4, 'cm'),
          plot.margin = unit(c(0, -2, 0, -2), "cm"),
          legend.position = "bottom",
          legend.box.just = "left",
          legend.box = "vertical")

ggsave("figures/figure1_study_area.png",
       bg = "white",
       width = 3.5, height = 4.5, dpi = 500)
```

## Results
### Figure 2

```{r}
f_dir <- "results/scenarios"
items <- c("Y", "B", "Ca", "Co", "D", "hybrid")
simu_mean <- do.call(rbind, lapply(items, function(item){
    read.csv(sprintf("%s/%s_simulation_mean.csv", f_dir, item)) %>% 
        mutate(mod = item)
})) %>% mutate(mod = factor(mod, levels = c("Y", "B", "Ca", "Co", "D", "hybrid"),
                            labels = c("Y", "B", "Ca", "Co", "D", "Hybrid")))

simu_sd <- do.call(rbind, lapply(items, function(item){
    read.csv(sprintf("%s/%s_simulation_sd.csv", f_dir, item)) %>% 
        mutate(mod = item)
})) %>% mutate(mod = factor(mod, levels = c("Y", "B", "Ca", "Co", "D", "hybrid"),
                            labels = c("Y", "B", "Ca", "Co", "D", "Hybrid")))

prod_gain_intense <- read.csv(
    file.path(tdf_dir, "production_gain_from_intensifify.csv"))
pgi_mean <- mean(prod_gain_intense$Production_gain) / 1e7
pgi_sd <- sd(prod_gain_intense$Production_gain) / 1e7

fig_list <- lapply(names(simu_mean)[2:6], function(ele){
    vals <- simu_mean %>% select(all_of(c("farmable_ratio", "mod", ele))) %>% 
        mutate(sd = simu_sd %>% 
                   select(all_of(c("farmable_ratio", "mod", ele))) %>%
                   pull(all_of(ele)))
    names(vals)[3:4] <- c("value", "sd")
    vals$farmable_ratio <- vals$farmable_ratio / 1e6
    
    # Separate different conditions
    if (str_detect(ele, "Production")){
        vals$value <- vals$value / 1e7 + pgi_mean
        vals$sd <- vals$sd / 1e7
        ylab_nm <- expression(paste("Production gain (tons * ", 10^7, ")"))
    } else if (str_detect(ele, "Biodiversity")) {
        vals$value <- vals$value / 1e5
        vals$sd <- vals$sd / 1e5
        ylab_nm <- expression(paste("Biodiversity loss (", 10^5, ")"))
    } else if (str_detect(ele, "Carbon")){
        vals$value <- vals$value / 1e7
        vals$sd <- vals$sd / 1e7
        ylab_nm <- expression(paste("Carbon loss (tons * ", 10^7, ")"))
    } else if (str_detect(ele, "Connectivity")){
        vals$value <- vals$value / 1e5
        vals$sd <- vals$sd / 1e5
        ylab_nm <- expression(paste("Connectivity loss (", 10^5, ")"))
    } else {
        vals$value <- vals$value / 1e8
        vals$sd <- vals$sd / 1e8
        ylab_nm <- expression(paste("Cost distance increase (", 10^8, ")"))
    }
    
    p <- ggplot(vals, aes(x = farmable_ratio, y = value)) + 
        geom_ribbon(aes(ymin = value - sd, ymax = value + sd, fill = mod), 
                    alpha = 0.2) +
        geom_line(aes(color = mod)) + 
        scale_color_manual(
            "Optimize to", values = c("#E69F00", "#009E73", "#D55E00", 
                                   "#0072B2", "#999999", "#000000"),
            labels = c("Yield", "Biodiversity", "Carbon", "Connectivity", 
                       "Cost distance", "Hybrid"), 
            aesthetics = c("colour", "fill"))
    
    if (str_detect(ele, "Production")){
        p <- p + geom_point(aes(x = min(farmable_ratio), y = pgi_mean), size = 0.8) +
            geom_errorbar(aes(x = min(farmable_ratio), y = pgi_mean,
                              ymin = pgi_mean - pgi_sd, ymax = pgi_mean + pgi_sd), 
                          width = .2, position = position_dodge(0.05)) + 
            annotate(geom = "text", x = 0.25, hjust = 0, size = 2.1,
                     y = pgi_mean, label = "Production gain from existing cropland")
    }
    
    p + xlab(expression(paste("Expansion in cropland (ha * ", 10^6, ")"))) +
        ylab(ylab_nm) + theme_bw() +
        theme(text = element_text(size = 10),
              legend.text = element_text(size = 12),
              legend.title = element_text(size = 12),
              legend.key.height = unit(0.8, 'cm'),
              legend.key.width = unit(0.8, 'cm'),
              legend.position = "none",
              plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"))
})

legend <- get_legend(fig_list[[1]] + theme(legend.position = "right"))

ggarrange(fig_list[[1]], fig_list[[2]], fig_list[[3]], 
          fig_list[[4]], fig_list[[5]], legend, nrow = 2, ncol = 3,
          labels = c("A", "B", "C", "D", "E", ""))

ggsave("figures/figure2_scenarios.png",
       bg = "white",
       width = 7.4, height = 5.5, dpi = 500)
```

### Figure 3

```{r}
f_dir <- "results/compare"
items <- c("RG", "HY", "CS")

prod_gain_intense <- read.csv(
    file.path(tdf_dir, "production_gain_from_intensifify.csv"))
pgi_mean_rg <- mean(prod_gain_intense$Production_gain) / 1e7
pgi_sd_rg <- sd(prod_gain_intense$Production_gain) / 1e7

prod_gain_intense <- read.csv(
    file.path(tdf_dir, "production_gain_from_intensifify_150.csv"))
pgi_mean_hy <- mean(prod_gain_intense$Production_gain) / 1e7
pgi_sd_hy <- sd(prod_gain_intense$Production_gain) / 1e7

prod_gain_intense <- read.csv(
    file.path(tdf_dir, "production_gain_from_intensifify_more_cassava.csv"))
pgi_mean_cs <- mean(prod_gain_intense$Production_gain) / 1e7
pgi_sd_cs <- sd(prod_gain_intense$Production_gain) / 1e7

simu_mean <- do.call(rbind, lapply(items, function(item){
    tb <- read.csv(sprintf("%s/%s_hybrid_mean.csv", f_dir, item)) %>% 
        mutate(mod = item)
    if (item == "RG"){
        tb %>% mutate(Production_gain = Production_gain / 1e7 + pgi_mean_rg)
    } else if (item == "HY"){
        tb %>% mutate(Production_gain = Production_gain / 1e7 + pgi_mean_hy)
    } else {
        tb %>% mutate(Production_gain = Production_gain / 1e7 + pgi_mean_cs)
    }
})) %>% mutate(mod = factor(mod, levels = items, labels = items))

simu_sd <- do.call(rbind, lapply(items, function(item){
    read.csv(sprintf("%s/%s_hybrid_sd.csv", f_dir, item)) %>% 
        mutate(mod = item) %>% 
        mutate(Production_gain = Production_gain / 1e7)
})) %>% mutate(mod = factor(mod, levels = items, labels = items))

vals <- simu_mean %>% select(c("farmable_ratio", "mod", "Production_gain")) %>% 
        mutate(sd = simu_sd %>% pull("Production_gain"))
vals$farmable_ratio <- vals$farmable_ratio / 1e6

p1 <- ggplot(vals, aes(x = farmable_ratio, y = Production_gain)) + 
    geom_ribbon(aes(ymin = Production_gain - sd, ymax = Production_gain + sd, fill = mod), 
                alpha = 0.2) +
    geom_line(aes(color = mod)) + 
    scale_color_manual(
        "", values = c("#009E73", "#D55E00", "#0072B2"),
        labels = c("A", 'B', "C"), 
        aesthetics = c("colour", "fill")) +
    geom_point(aes(x = min(farmable_ratio), y = pgi_mean_rg), color = "#009E73", size = 0.8) +
    geom_point(aes(x = min(farmable_ratio), y = pgi_mean_hy), color = "#D55E00", size = 0.8) +
    geom_point(aes(x = min(farmable_ratio), y = pgi_mean_cs), color = "#0072B2", size = 0.8) +
    xlab(expression(paste("Expansion in cropland (ha * ", 10^6, ")"))) +
    ylab(expression(paste("Production gain (tons * ", 10^7, ")"))) + 
    theme_bw() +
    theme(text = element_text(size = 8),
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8),
          legend.key.height = unit(0.3, 'cm'),
          legend.key.width = unit(0.3, 'cm'),
          legend.margin = margin(0,0,0,0),
          legend.box.margin=margin(0,0,-5,0),
          legend.position = "top",
          plot.margin = unit(rep(0.1, 4), "cm"))

# Maps
f_dir <- "results/spatial"
items <- c("RG", "HY", "CS")

fig_list <- lapply(items, function(item){
    lyr <- rast(file.path(f_dir, sprintf("fut_agro_land_hybrid_%s.tif", item)))
    lyr[is.na(lyr)] <- 0
    lyr <- (lyr$maize > 95) * 1 + (lyr$paddy > 95) * 2 + (lyr$sorghum > 95) * 3 +
        (lyr$cassava > 95) * 4 + (lyr$beans > 95) * 5
    lyr[lyr == 0] <- NA
    
    cls <- data.frame(id = 1:5, cover = c("Maize", "Rice", "Sorghum", "Cassava", "Beans"))
    levels(lyr) <- cls
    
    ggplot() +
        geom_spatraster(data = lyr, na.rm = TRUE) +
        scale_fill_manual(
            name = "Suggested crop", na.translate = FALSE,
            values = c("#E69F00", "#009E73", "#D55E00", "#0072B2", "darkgrey")) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.4) +
        coord_sf() + theme_void() +
        theme(text = element_text(size = 8),
              legend.text = element_text(size = 8),
              plot.margin = unit(c(0, 0, 0, 0), "cm"))
})

ggarrange(
    p1, ggarrange(fig_list[[1]], fig_list[[2]], fig_list[[3]], nrow = 1, ncol = 3,
                  labels = c("A", "B", "C"), hjust = -4, 
                  common.legend = TRUE, legend = "right"),
    nrow = 1, ncol = 2, widths = c(1, 4))

ggsave("figures/figure3_practices.png",
       bg = "white",
       width = 7.4, height = 2, dpi = 500)
```

### Figure 3

To show the significance of productivity growth on existing cropland.

```{r}
fnames <- list.files("results/intensification", full.names = TRUE)
vals <- do.call(rbind, lapply(fnames, function(fname){
    params <- as.numeric(str_extract_all(fname, "[0-9]+")[[1]])
    vals <- read.csv(fname) %>% pull(Production_gain)
    data.frame(cassava = params[1], yield = params[2], 
               land_usage = params[3], mean = mean(vals),
               sd = sd(vals))
}))

# Figure for cassava increase
g1 <- ggplot(data = vals %>% filter(yield == 100),
       aes(y = mean/1e7, x=cassava, color = factor(land_usage))) +
    geom_hline(yintercept = 12572439/1e7, linetype = "dashed", color = "grey") + 
    geom_point() + geom_line() + ylim(0, max(vals$mean)/1e7) +
    annotate(geom = "text", x = 3, size = 2.5,
                     y = 12002439/1e7, label = "Double the current production") +
    scale_colour_brewer("Land usage (%)", palette = "Dark2") +
    geom_errorbar(aes(ymin = (mean - sd)/1e7, ymax = (mean + sd)/1e7), width=.1) +
    xlab("Increase (fold) in planted area of cassava") +
    ylab(expression(paste("Production growth (", 10^7, "tons)"))) +
    theme_classic()+
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.key.height = unit(0.3, 'cm'),
          legend.key.width = unit(0.3, 'cm'),
          legend.margin = margin(rep(0.1, 4)),
          legend.box.margin = margin(rep(0.2, 4)),
          legend.position = "top",
          plot.margin = unit(rep(0.1, 4), "cm"))

g2 <- ggplot(data = vals %>% filter(cassava == 1) %>% filter(yield < 150),
       aes(y = mean/1e7, x=yield, color = factor(land_usage))) +
    geom_hline(yintercept = 12572439/1e7, linetype = "dashed", color = "grey") + 
    geom_point() + geom_line() + ylim(0, max(vals$mean)/1e7) +
    annotate(geom = "text", x = 120, size = 2.5,
                     y = 12002439/1e7, label = "Double the current production") +
    scale_colour_brewer("Land usage (%)", palette = "Dark2") +
    geom_errorbar(aes(ymin = (mean - sd)/1e7, ymax = (mean + sd)/1e7), width=.1) +
    xlab("Increase (%) in yield gap close") +
    ylab(expression(paste("Production growth (", 10^7, "tons)"))) +
    theme_classic()+
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.key.height = unit(0.3, 'cm'),
          legend.key.width = unit(0.3, 'cm'),
          legend.margin = margin(rep(0.1, 4)),
          legend.box.margin = margin(rep(0.2, 4)),
          legend.position = "top",
          plot.margin = unit(rep(0.1, 4), "cm"))

ggarrange(g1, g2, nrow = 1, ncol = 2,
          labels = c("A", "B"), legend = "top",
          common.legend = TRUE)

ggsave("figures/figure3_prod_growth.png",
       bg = "white",
       width = 6, height = 3, dpi = 500)
```

### Figure 4

```{r}
fnames <- list.files("results/expansion", full.names = TRUE)
vals <- do.call(rbind, lapply(fnames, function(fname){
    params <- as.numeric(str_extract_all(fname, "[0-9]+")[[1]])
    vals <- read.csv(fname) %>% pull(Attainable_production)
    data.frame(cassava = params[1], yield = params[2], 
               land_usage = params[3], mean = mean(vals),
               sd = sd(vals))
}))

# Figure
# geom_bar(stat="identity", position=position_dodge()) + 
# Figure for cassava increase
g1 <- ggplot(data = vals %>% filter(yield == 100),
       aes(y = mean/1e7, x=cassava, color = factor(land_usage))) +
    geom_point() + geom_line() + ylim(min(vals$mean)/1e7, max(vals$mean)/1e7) +
    scale_colour_brewer("Land usage (%)", palette = "Dark2") +
    geom_errorbar(aes(ymin = (mean - sd)/1e7, ymax = (mean + sd)/1e7), width=.1) +
    xlab("Increase (fold) in planted area of cassava") +
    ylab(expression(paste("Production potential (", 10^7, "tons)"))) +
    theme_classic()+
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.key.height = unit(0.3, 'cm'),
          legend.key.width = unit(0.3, 'cm'),
          legend.margin = margin(rep(0.1, 4)),
          legend.box.margin = margin(rep(0.2, 4)),
          legend.position = "top",
          plot.margin = unit(rep(0.1, 4), "cm"))

g2 <- ggplot(data = vals %>% filter(cassava == 1) %>% filter(yield < 150),
       aes(y = mean/1e7, x=yield, color = factor(land_usage))) +
    geom_point() + geom_line() + ylim(min(vals$mean)/1e7, max(vals$mean)/1e7) +
    scale_colour_brewer("Land usage (%)", palette = "Dark2") +
    geom_errorbar(aes(ymin = (mean - sd)/1e7, ymax = (mean + sd)/1e7), width=.1) +
    xlab("Increase (%) in yield gap close") +
    ylab(expression(paste("Production potential (", 10^7, "tons)"))) +
    theme_classic()+
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.key.height = unit(0.3, 'cm'),
          legend.key.width = unit(0.3, 'cm'),
          legend.margin = margin(rep(0.1, 4)),
          legend.box.margin = margin(rep(0.2, 4)),
          legend.position = "top",
          plot.margin = unit(rep(0.1, 4), "cm"))

ggarrange(g1, g2, nrow = 1, ncol = 2,
          labels = c("A", "B"), legend = "top",
          common.legend = TRUE)

ggsave("figures/figure4_prod_potential.png",
       bg = "white",
       width = 6, height = 3, dpi = 500)
```

## Agricultural yield

### Figure S4

Down-scaled and calibrated current yield map of maize (A), rice (B), sorghum (C), cassava (D), and beans (E) at 1 km.

```{r}
fname <- file.path(agro_dir, "yield_calibrated_5crops_tz_1km.tif")
yields_calib <- rast(fname)

org_yields <- rast(file.path(agro_dir, "yields_10km.tif"))
names(org_yields) <- names(yields_calib)

fig_list <- lapply(names(yields_calib), function(nm){
    org_yield <- org_yields[[nm]]
    yield_calib <- yields_calib[[nm]]
    
    org_figure <- ggplot() +
        geom_spatraster(data = org_yield, na.rm = TRUE) +
        scale_fill_continuous_sequential(
            name = "Yield (t/ha)",
            palette = "Plasma", na.value = NA) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.4) +
        coord_sf() + theme_void() +
        theme(legend.position = "null")
    
    org_figure <- ggplotGrob(org_figure)
    
    ggplot() +
        geom_spatraster(data = yield_calib, na.rm = TRUE) +
        scale_fill_continuous_sequential(
            name = "Yield (t/ha)",
            palette = "Plasma", na.value = NA,
            guide = guide_colourbar(
                title.position = "top", title.hjust = 0.5)) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.6) +
        annotation_custom(grob = org_figure, 
                      xmin = 37.2, xmax = 40.6,
                      ymin = - 4.0, ymax = -0.8) +
        coord_sf() + theme_void() +
        theme(text = element_text(size = 12),
              strip.text.x = element_text(size = 12),
              legend.text = element_text(size = 12),
              legend.spacing.y = unit(0.1, 'cm'),
              legend.margin = margin(rep(0, 4)), 
              legend.key.height = unit(0.4, 'cm'),
              legend.key.width = unit(0.8, 'cm'),
              plot.margin = unit(c(0, 0, 0, 0), "cm"),
              legend.position = "bottom",
              legend.box = "vertical")
})
ggarrange(plotlist = fig_list, nrow = 2, ncol = 3,
          labels = c("A", "B", "C", "D", "E"))

ggsave("figures/S4_current_yields.png",
       bg = "white",
       width = 7.5, height = 5.5, dpi = 500)
```

### Table S1

```{r}
crops <- c("MAIZ", "RICE", "SORG", "CASS", "BEAN")
var_summary <- lapply(crops, function(crp){
    load(file.path(agro_dir, sprintf("qrf_%s.rda", crp)))
    variables <- rf_model$forest$independent.variable.names
    
    # Group the variables
    temperature <- variables[str_detect(variables, "tmean|tmax|tmin|gdd|srad[0-9]+")]
    temperature <- gsub("tm", "TM", temperature)
    temperature <- gsub("gdd", "GDD", temperature)
    temperature <- gsub("srad", "Srad", temperature)
    precipitation <- variables[str_detect(variables, "prec|vapr|et[0-9]+")]
    precipitation <- gsub("et", "ET", precipitation)
    precipitation <- gsub("prec", "Prec", precipitation)
    precipitation <- gsub("vapr", "Vapr", precipitation)
    soil <- variables[str_detect(variables, "_mean")]
    soil <- gsub("_mean", "", soil)
    soil <- gsub("\\.", "-", soil)
    
    # Add them together to a table
    ## temp
    temperature <- sapply(c("TMean", "TMax", "TMin", "GDD", "Srad"), function(item){
        item_list <- temperature[grep(item, temperature)]
        paste(item_list, collapse = ", ")
    }) %>% data.frame(Selected = .) %>% mutate(Group = row.names(.)) %>% 
        mutate(Element = "Temperature")
    
    ## prec
    precipitation <- sapply(c("ET", "Prec", "Vapr"), function(item){
        item_list <- precipitation[grep(item, precipitation)]
        paste(item_list, collapse = ", ")
    }) %>% data.frame(Selected = .) %>% mutate(Group = row.names(.)) %>% 
        mutate(Element = "Precipitation")
    
    ## soil
    soil <- sapply(c("bdod", "cec", "cfvo", "clay", 
                     "nitrogen", "phh2o", "sand", "silt", "soc"), function(item){
        item_list <- soil[grep(item, soil)]
        paste(item_list, collapse = ", ")
    }) %>% data.frame(Selected = .) %>% mutate(Group = row.names(.)) %>% 
        mutate(Element = "Soil")
    
    tb <- rbind(temperature, precipitation, soil) %>%
        select(Element, Group, Selected)
    names(tb)[3] <- crp
    tb
})

var_summary <- Reduce(full_join, var_summary)
write.csv(var_summary, "figures/table_s1_selected_vars.csv", row.names = FALSE)
```

### Figure S5

Attainable yield (predicted 95% quantile of crop yield downscaling model) map at 1 km.

```{r}
fname <- file.path(agro_dir, "yield_attainable_5crops_tz_1km.tif")
yields_atn <- rast(fname)

fig_list <- lapply(names(yields_atn), function(nm){
    yield_atn <- yields_atn[[nm]]
    
    ggplot() +
        geom_spatraster(data = yield_atn, na.rm = TRUE) +
        scale_fill_continuous_sequential(
            name = "Yield (t/ha)",
            palette = "Plasma", na.value = NA,
            guide = guide_colourbar(
                title.position = "top", title.hjust = 0.5)) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.6) +
        coord_sf() + theme_void() +
        theme(text = element_text(size = 12),
              strip.text.x = element_text(size = 12),
              legend.text = element_text(size = 12),
              legend.spacing.y = unit(0.1, 'cm'),
              legend.margin = margin(rep(0, 4)), 
              legend.key.height = unit(0.4, 'cm'),
              legend.key.width = unit(0.8, 'cm'),
              plot.margin = unit(c(0, 0, 0, 0), "cm"),
              legend.position = "bottom",
              legend.box = "vertical")
})
ggarrange(plotlist = fig_list, nrow = 2, ncol = 3,
          labels = c("A", "B", "C", "D", "E"))

ggsave("figures/S5_attainable_yields.png",
       bg = "white",
       width = 7.5, height = 5.5, dpi = 500)
```

## Refine species habitat type map and range map

### Figure S6

Habitat type map of Tanzania before and after refining.

```{r}
hbt_dir <- "data/habitats"
habitat_lv2 <- rast(file.path(hbt_dir, "habitat_tz_lvl2.tif"))
habitat_refined <- rast(file.path(hbt_dir, "habitat_tz_refined_final.tif"))

# Plot
clr <- read.delim2(file.path(hbt_dir, "styles/level2.clr"), header = FALSE)[[1]]
clr <- do.call(rbind, lapply(clr, function(r){
    nums <- as.integer(str_extract_all(r, "[0-9]+ ")[[1]])[1:5]
    nm <- gsub(paste(nums, collapse = " "), "", r)
    nm <- gsub("^ ", "", nm)
    
    matrix(c(nums[1:4], nm), nrow = 1, ncol = 5)
}))

levels(habitat_lv2) <- levels(habitat_refined) <- 
    data.frame(value = as.integer(clr[, 1]), landcover = clr[, 5])
coltab(habitat_lv2) <- coltab(habitat_refined) <- 
    data.frame(value = as.integer(clr[, 1]), red = as.integer(clr[, 2]), 
               blue = as.integer(clr[, 3]), green = as.integer(clr[, 4]))

# Plot
cols <- coltab(habitat_lv2)[[1]]
cols <- sapply(1:nrow(cols), function(i){
    rgb(cols[i, 2:4], maxColorValue = 255)})

# Can't believe I have to rely on base R plot for efficient plotting in 2023.
p1 <- function() {
    par(mar = rep(0, 4))
    plot(habitat_lv2, axes = FALSE, box = FALSE, legend = FALSE)}
p2 <- function() {
    par(mar = rep(0, 4))
    plot(habitat_refined, axes = FALSE, box = FALSE, legend = FALSE)}

legend_lc <- ggplot() + 
    geom_raster(
        data = data.frame(rst = as.factor(coltab(habitat_lv2)[[1]][, 1]),
                          x = 1:length(cols),
                          y = 1:length(cols)), 
        aes(x = x, y = y, 
            fill = rst)) +
    scale_fill_manual('',
                      values = cols, 
                      breaks = coltab(habitat_lv2)[[1]][, 1],
                      labels = clr[, 5]) +
    theme(legend.position = 'bottom',
          text = element_text(size = 8),
          legend.text = element_text(size = 8))+
    guides(fill = guide_legend(ncol = 2))
legend_lc <- get_legend(legend_lc)

ggarrange(NULL, ggarrange(ggdraw(p1), ggdraw(p2), nrow = 1, ncol = 2,
          labels = c("A", "B"), vjust=6), NULL, legend_lc, NULL, nrow = 5, 
          heights = c(-0.4, 4, 2.5, 0.5, 3))

ggsave("figures/S6_habitat_type.png",
       bg = "white",
       width = 8, height = 11.2, dpi = 500)
```

## Landscape connectivity

### Table S2

Selected species to estimate landscape connectivity. This is not the table itself, but get all necessary elements to generate Table S2.

```{r}
conn_dir <- "data/connectivity"
selected_species <- read.csv(file.path(conn_dir, "select_species.csv"))
sdm_eval <- read.csv(file.path(conn_dir, "multi_species_sdm_eval.csv"))
tb <- full_join(selected_species, sdm_eval, by = c("sci_name" = "species")) %>% 
    select(sci_name, category, pup_trend, KAPPA, TSS, ROC)

vars <- rast(file.path(conn_dir, "variables/variables.tif")) %>% names()
```

### Figure S7

```{r}
suitability <- rast(file.path(conn_dir, "circuit/suitability.asc"))
pas <- read_sf(file.path(conn_dir, "circuit/pas_selected.geojson")) %>% select()

ggplot() +
    geom_spatraster(data = suitability, na.rm = TRUE) +
    scale_fill_continuous_sequential(
            name = "Suitability (0-1)",
            palette = "Plasma", na.value = NA,
            guide = guide_colourbar(
                title.position = "top", title.hjust = 0.5)) +
    geom_sf(data = pas, color = "white", show.legend = FALSE,
            fill = "transparent", linewidth = 0.4) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.6) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.text = element_text(size = 10),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.8, 'cm'),
          plot.margin = unit(c(-1, -0.5, 0, -0.5), "cm"),
          legend.position = "bottom",
          legend.box.just = "left",
          legend.box = "vertical")

ggsave("figures/S7_species_suitability.png",
       bg = "white", width = 3.5, height = 4.2, dpi = 500)
```

## Biodiversity index

### Table S3

Information of species to calculate biodiversity.

```{r}
bio_dir <- "data/biodiversity"
species_info <- read.csv(file.path(bio_dir, "species_info.csv")) %>% 
    filter(tanzania_range_size > 0)

# Numbers
species_info %>% group_by(taxon) %>% summarise(n = n())

# Endemic species numbers
species_info %>% filter(outside_range_size == 0) %>% 
    group_by(taxon) %>% summarise(n = n())

# Get median values
species_info %>% group_by(taxon) %>% 
    summarise(median = median(refined_range_size))

species_info %>% group_by(taxon) %>% 
    summarise(median = median(range_size))
```

### Figure S8

Weighted species richness and rarity-weighted species richness.

```{r}
# Only use weighted
fnames <- list.files(bio_dir, pattern = "*weighted*", full.names = TRUE)

# Calculate species component
normalize <- function(x, robust = TRUE) {
    if (robust) {
        stretch(x, minv = 0, maxv = 1, minq = 0.01, maxq = 0.99)
    } else {
        (x - minmax(x)[1]) / (minmax(x)[2] - minmax(x)[1])
    }
}

# Richness
richness_fns <- fnames[!grepl("rarity", fnames)]
richness_layers <- do.call(
    c, lapply(richness_fns, function(fn){
        normalize(rast(fn) %>% mask(bry))}))
names(richness_layers) <- c("Amphibians", "Birds", "Mammals", "Reptiles")
richness <- richness_layers %>% mean() %>% normalize(FALSE)

endemism_fns <- fnames[grepl("rarity", fnames)]
# Use robust scaling to avoid the impacts of outliers
endemism_layers <- do.call(
    c, lapply(endemism_fns, function(fn){
        normalize(sqrt(sqrt(rast(fn) %>% mask(bry))))}))
names(endemism_layers) <- c("Amphibians", "Birds", "Mammals", "Reptiles")
endemism <- endemism_layers %>% mean() %>% normalize(FALSE)

## Plot
p1 <- ggplot() +
    geom_spatraster(data = richness, na.rm = TRUE) +
    scale_fill_continuous_sequential(
        name = "Normalized value (0-1)",
        palette = "Plasma", na.value = NA,
        guide = guide_colourbar(
            title.position = "top", title.hjust = 0.5)) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.6) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 12),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size = 12),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.8, 'cm'),
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          legend.position = "bottom",
          legend.box = "vertical")

fig_list <- lapply(c("Mammals", "Birds", "Amphibians", "Reptiles"),
                   function(nm){
    richness_layer <- richness_layers[[nm]]
    
    ggplot() +
        geom_spatraster(data = richness_layer, na.rm = TRUE) +
        scale_fill_continuous_sequential(
            name = "Normalized value (0-1)",
            palette = "Plasma", na.value = NA,
            guide = guide_colourbar(
                title.position = "top", title.hjust = 0.5)) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.6) +
        ggtitle(nm) + 
        coord_sf() + theme_void() +
        theme(text = element_text(size = 12),
              plot.title = element_text(hjust = 0.5, face = "bold"),
              plot.margin = unit(c(0, 0, 0, 0), "cm"),
              legend.position = "none")
})
p2 <- ggarrange(plotlist = fig_list, nrow = 2, ncol = 2)

p3 <- ggplot() +
    geom_spatraster(data = endemism, na.rm = TRUE) +
    scale_fill_continuous_sequential(
        name = "Normalized value (0-1)",
        palette = "Plasma", na.value = NA,
        guide = guide_colourbar(
            title.position = "top", title.hjust = 0.5)) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.6) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 12),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size = 12),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.8, 'cm'),
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          legend.position = "bottom",
          legend.box = "vertical")

fig_list <- lapply(c("Mammals", "Birds", "Amphibians", "Reptiles"),
                   function(nm){
    endemism_layer <- endemism_layers[[nm]]
    
    ggplot() +
        geom_spatraster(data = endemism_layer, na.rm = TRUE) +
        scale_fill_continuous_sequential(
            name = "Normalized value (0-1)",
            palette = "Plasma", na.value = NA,
            guide = guide_colourbar(
                title.position = "top", title.hjust = 0.5)) +
        geom_sf(data = bry, color = "black", 
                fill = "transparent", linewidth = 0.6) +
        ggtitle(nm) + 
        coord_sf() + theme_void() +
        theme(text = element_text(size = 12),
              plot.title = element_text(hjust = 0.5, face = "bold"),
              plot.margin = unit(c(0, 0, 0, 0), "cm"),
              legend.position = "none")
})
p4 <- ggarrange(plotlist = fig_list, nrow = 2, ncol = 2)

g1 <- ggarrange(p1 + theme(legend.position = "none"), p2, nrow = 1, ncol = 2) +
    theme(panel.background = element_rect(fill = "lightgrey", color = "lightgrey"))

g2 <- ggarrange(p3 + theme(legend.position = "none"), p4, nrow = 1, ncol = 2)

ggarrange(g1, g2, get_legend(p1), nrow = 3, ncol = 1,
          labels = c("A", "B", ""), heights = c(1, 1, 0.15))

ggsave("figures/S8_species_richness.png",
       bg = "white",
       width = 7.5, height = 7.6, dpi = 500)
```

### Figure S9

```{r}
bhi_msa_bii <- rast(file.path(bio_dir, "bhi_msa_bii.tif"))
bhi_msa_bii <- mask(bhi_msa_bii, bry)

fig_list <- lapply(names(bhi_msa_bii), function(nm){
    lyr <- bhi_msa_bii[[nm]]
    
    ggplot() +
    geom_spatraster(data = lyr, na.rm = TRUE) +
    scale_fill_continuous_sequential(
        name = "Value (0-1)",
        palette = "Plasma", na.value = NA,
        guide = guide_colourbar(
            title.position = "top", title.hjust = 0.5)) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.6) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 12),
          strip.text.x = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
          legend.text = element_text(size = 12),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.8, 'cm'),
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          legend.position = "bottom",
          legend.box = "vertical")
})

ggarrange(plotlist = fig_list, nrow = 1, ncol = 3,
          labels = c("A", "B", "C"))

ggsave("figures/S9_ecological_index.png",
       bg = "white",
       width = 7, height = 3, dpi = 500)
```

### Figure S10

```{r}
carbon_dir <- "data/carbon"
abv_bms <- rast(file.path(carbon_dir, "aboveground_biomass.tif"))
blw_bms <- rast(file.path(carbon_dir, "belowground_biomass.tif"))
soct_1m <- rast(file.path(carbon_dir, "ocs_1m_sg.tif")) %>% 
    resample(abv_bms)

carbons <- c(abv_bms, blw_bms, soct_1m)
names(carbons) <- c("Aboveground biomass", "Belowground biomass", 
                    "Soil carbon stocks")
carbons <- mask(carbons, bry)

fig_list <- lapply(names(carbons), function(nm){
    lyr <- carbons[[nm]]
    
    ggplot() +
    geom_spatraster(data = lyr, na.rm = TRUE) +
    scale_fill_continuous_sequential(
        name = "Carbon density (t/ha)",
        palette = "Plasma", na.value = NA,
        guide = guide_colourbar(
            title.position = "top", title.hjust = 0.5)) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.6) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 12),
          strip.text.x = element_text(size = 12),
          plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
          legend.text = element_text(size = 12),
          legend.spacing.y = unit(0.1, 'cm'),
          legend.margin = margin(rep(0, 4)), 
          legend.key.height = unit(0.4, 'cm'),
          legend.key.width = unit(0.8, 'cm'),
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          legend.position = "bottom",
          legend.box = "vertical")
})

ggarrange(plotlist = fig_list, nrow = 1, ncol = 3,
          labels = c("A", "B", "C"))

ggsave("figures/S10_carbons.png",
       bg = "white",
       width = 7, height = 3, dpi = 500)
```

