---
title: 'Part1: agriculture analysis'
author: "Lei Song"
date: "1/4/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(sf)
library(here)
library(dplyr)
library(terra)
library(stringr)
library(purrr)
library(rcropmod)
library(Dasst)
select <- dplyr::select

# Set paths
dssat_data_dir <- "/Users/pinot/Dropbox/dssat48"
dssat_exe <- file.path(dssat_data_dir, "dscsm048")
soil_dir <- file.path(dssat_data_dir, "Soil")
weather_dir <- file.path(dssat_data_dir, "Weather")
geno_dir <- file.path(dssat_data_dir, "Genotype")
agro_dir <- file.path(here("data/agriculture"))

result_dir <- here("results")
agro_result <- file.path(result_dir, "agriculture")
```

## Introduction

This document covers the analysis of agriculture:

- Current crop production at current farmland
- Crop production gap at current farmland
- Farming potential at available areas for possible expansion.

We focus on two staple crops: Maize and Paddy.

The main methods are:

- Run DSSAT using different management strategies for simulation based on the reality in Tanzania.
- Run statistical model to predict gridded crop production and gaps.

## Soil classification

The weather station data is not sufficient to run DSSAT, we won't use it for our simulation, so we don't have to stick with †he stations. Thus, we regularly sampled points across the whole country to run the model. The soil classification map will be used to collect samples for DSSAT yield estimation. We queried the map from SoilGrids website (https://soilgrids.org) hosted by ISRIC.

```{r}
bry <- read_sf(here("data/geoms/mainland_tanzania.geojson")) %>% 
    select()

coords <- st_bbox(bry)
coords <- sprintf("long(%s,%s)&SUBSET=lat(%s,%s)", 
                  coords[1] - 0.2, coords[3] + 0.2,
                  coords[2] - 0.2, coords[4] + 0.2)

url <- sprintf("https://maps.isric.org/mapserv?map=/map/wrb.map&SERVICE=WCS&VERSION=2.0.1&REQUEST=GetCoverage&COVERAGEID=MostProbable&FORMAT=image/tiff&SUBSET=%s&SUBSETTINGCRS=http://www.opengis.net/def/crs/EPSG/0/4326&OUTPUTCRS=http://www.opengis.net/def/crs/EPSG/0/4326", coords)

download.file(url = url, destfile = here("data/soil/WRD_soil_class.tif"))
```

## Prepare samples for simulation

We randomly sampled points for each soil type within each district to run DSSAT simulation. We only used the cropland area (with a reference of year 2018).

```{r}
# Read districts
tz_l1 <- read_sf(here("data/geoms/tanzania_level1.geojson")) %>% 
    mutate(id = as.integer(gsub("TZ", "", ADM1_PCODE))) %>% 
    select(id) %>% 
    st_make_valid()

# Read Cropland
crop_mask <- rast(here("data/agriculture/landcover_1km.tif"))
crop_mask[crop_mask != 1] <- NA

# Subset soil layer
soil_class <- rast(here("data/soil/WRD_soil_class.tif"))
tz_l1 <- rasterize(vect(tz_l1), soil_class, field = "id")
crop_mask <- project(crop_mask, soil_class, method = "near")
crop_mask <- resample(crop_mask, soil_class, method = "near")

soil_class <- c(soil_class, tz_l1) %>% 
    mask(crop_mask) %>% 
    mask(vect(bry))

# Sample the points
## First, summarize the maps
soil_df <- values(soil_class) %>% na.omit() %>% 
    data.frame() %>% 
    rename(district = id,
           soil_type = WRD_soil_class)
soil_type_l1 <- soil_df %>% 
    group_by(district) %>% 
    summarise(n = length(unique(soil_type)))

soil_pts <- read_sf(
    file.path("data/soil/han",
              "point5m_soilgrids-for-dssat-10km_v1.shp")) %>% 
    st_intersection(bry)

soil_classes <- extract(soil_class, soil_pts)
soil_pts <- soil_pts %>% 
    mutate(soil_class = soil_classes$WRD_soil_class,
           adm1_pcode = soil_classes$id) %>% na.omit()
soil_pts <- soil_pts %>% st_drop_geometry() %>% 
    select(adm1_pcode, soil_class, SoilProfil, X, Y)

set.seed(1234)
samples <- soil_pts %>% 
    group_by(adm1_pcode, soil_class) %>% 
    sample_n(size = 1) %>% 
    rename(lat = Y, lon = X) %>% 
    ungroup() %>% 
    mutate(id = 1:nrow(.),
           WeatherProfil = paste0("TZA", 1:nrow(.) + 10000),
           adm1_pcode = paste0("TZ", str_pad(adm1_pcode, 2, pad = "0"))) %>% 
    select(id, adm1_pcode, WeatherProfil, SoilProfil, lat, lon)
write.csv(samples, here("data/geoms/dssat_samples.csv"), row.names = FALSE)
```

## Weather dataset

We queried meteorological and solar data from NASA POWER and CHIRPS. First, we made the input file using the selected samples, and then queried the data using python scripts shared by [NASAP_CHIRPS_WTH](https://github.com/DSSAT/NASAP_CHIRPS_WTH).

```{r}
stations_nasa <- samples %>% 
    select(id, SoilProfil, lat, lon) %>% 
    rename(nasapid = id, 
           Latitude = lat, Longitude = lon) %>% 
    mutate(ID = nasapid, LatNP = Latitude, LonNP = Longitude) %>% 
    select(ID, Latitude, Longitude, nasapid, LatNP, LonNP)
write.csv(stations_nasa, here("data/weather/nasap_query.csv"), 
          row.names = FALSE)
```

Then within terminal

```
python NASAP_CHIRPS_WTH get /path/to/nasap_query.csv 20110101 20201231 /path/to/comTCA/data/weather/wth
```

```{r}
# Rename the files to fit with DSSAT tradition
# Because NASAP_CHIRPS_WTH only takes numeric index
walk(list.files(here("data/weather/wth")), function(fname){
    new_fname <- paste0(
        "TZA", as.integer(tools::file_path_sans_ext(fname)) + 10000, ".WTH")
    file.rename(file.path(here("data/weather/wth"), fname), 
                file.path(here("data/weather/wth"), new_fname))
})
```

In terminal: `cp -R /path/to/comTCA/data/weather/wth/* weather_dir` to copy weather files into DSSAT Data directory.

## Soil dataset

We queried SoilGrid 10km properties made by [Han et al.](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/1PEEY0) to be the soil used by DSSAT. We copied the TZ.SOL to DSSAT working directory, and the soil-weather grid index catalog is `data/geoms/dssat_samples.csv`.

```{r}
# Back-up existing file
file.rename(file.path(soil_dir, "SOIL.SOL"), 
            file.path(soil_dir, "SOIL.BAK"))

# Copy new one over
file.copy(file.path("data/soil/han", "TZ.SOL"),
          file.path(soil_dir, "SOIL.SOL"))
```

## Crop calendar

For Tanzania, agricultural calendar starts from 1st October and ends on the 30th September of the subsequent year. Crop calendar was queried from [here](http://riskprofilesundrr.org/documents/1658).

**Maize**

- Long season: planting in Mar, and harvest in Jun to Sep.
- Short season: planting in Oct to Dec, and harvest in Dec to Feb.

**Rice**

Planting in Nov to Jan, and harvest in May to Jul.

## Run experiments
### Maize
#### Baseline (current yield)

District-level yield, seed usage, fertilizer usage, and planting dates are used to calibrate the model. The result is used as the baseline for maize production (current yield). Cultivation of maize is conducted in two different rainfall seasons of the year ([USDA](https://ipad.fas.usda.gov/rssiws/al/crop_calendar/safrica.aspx)):

- Bimodal (Vuli) October to December. This is the short rainy season.
- Unimodal (Msumi) March to May. This is the long rainy season.

under different nitrogen levels (0, 15, 30 and 60 kg N/ha) and rain-fed conditions. For the 15/30 kg N/ha treatment, DAP fertilizer was applied once after crop establishment, 28 days after sowing (DAS) whereas for the 60 Kg N/ha treatment, N fertilizer was applied in two rounds, the first one during planting to supply 30 Kg N/ha (as DAP) and the second round at 45 DAS as Urea, to supply the remaining 30 kg N/ha.

```{r}
districts <- read_sf(here("data/geoms/tanzania_level1.geojson")) %>% 
    mutate(Region = ifelse(
        ADM1_EN == "Dar-es-salaam", 
        "Dar Es Salaam", ADM1_EN),
        adm1_pcode = ADM1_PCODE) %>% 
    select(Region, adm1_pcode)

# See usage
source("scripts/query_seed_usage.R")
src_dir <- file.path(
    "data/agriculture/NSCA",
    "Small_Scale_Farmers_Dataset_ACQI")
seed_usage <- do.call(
    rbind, lapply(c("long", "short"), function(season){
        do.call(rbind, lapply(c("maize", "paddy"), function(crop){
            query_seed_usage(src_dir, season, crop)}))}))

# Irrigation usage
source("scripts/query_management.R")
irrigation_usage <- do.call(
    rbind, lapply(c("long", "short"), function(season){
        do.call(rbind, lapply(c("maize", "paddy"), function(crop){
            query_management(src_dir, season, crop, "irrigation")}))}))

# Fertilizer usage
fertilizer_usage <- do.call(
    rbind, lapply(c("long", "short"), function(season){
        do.call(rbind, lapply(c("maize", "paddy"), function(crop){
            query_management(src_dir, season, crop, "fertilizer")}))}))

# Burn to spatial data, take short-season maize as an example
short_maize <- seed_usage %>% 
    filter(Season == "short" & Crop == "maize") %>% 
    merge(districts, ., by = "Region")
write_sf(short_maize, "data/agriculture/short_maize.geojson")
```

Based on the surveys, we could tell that planting date, fertilizer usage, and probably planting management (plant density and row spacing) could be different across different districts. So we first tune these parameters using common maize cultivars in Tanzania.

##### Cultivar

Prepare cultivars and parameters for calibration with year 2019 - 2020:

```{r}
library(DSSAT)
source(here("scripts/read_write_cultivar.R"))

# Make the cultivar file
cul <- dssat_read_cultigen(file.path(geno_dir, "MZCER048.CUL"))
# cul <- cul[1:5, ]
# cul$`@VAR#` <- paste0("TZ", str_pad(1:5, 4, pad = "0"))
# cul$VRNAME.......... <- c("CM1509(LL)", "H614(LI)",  "CM1510(SL)", 
#                           "TMV1(SI)", "Staha(SI)")
# cul$`ECO#` <- rep("IB0001", 5)
# cul$P1 <- c(238.6, 290.8, 125, 215, 230.5)
# cul$P2 <- c(0.5, 0.471, 0.5, 0.5, 0.5)
# cul$P5 <- c(654, 921.2, 560, 635, 735)
# cul$G2 <- c(450, 796.8, 450, 650, 700)
# cul$G3 <- c(8.5, 5.26, 9.5, 7.55, 8.8)
# cul$PHINT <- c(75, 39.74, 75, 38, 47.3)
cul <- cul[1:2, ]
cul$`@VAR#` <- paste0("TZ", str_pad(1:2, 4, pad = "0"))
cul$VRNAME.......... <- c("SS1", "SS2")
cul$`ECO#` <- rep("IB0001", 2)
cul$P1 <- c(110, 125)
cul$P2 <- c(0.3, 0.5)
cul$P5 <- c(680, 560)
cul$G2 <- c(820, 450)
cul$G3 <- c(6.6, 9.5)
cul$PHINT <- c(38.9, 38.9)
dssatr_write_cultigen(
    cul, file.name = "MZCER048.CUL",
    dssat_path = dssat_data_dir,
    output_dir = here("data/agriculture"))

# Back-up existing file
file.rename(file.path(geno_dir, "MZCER048.CUL"), 
            file.path(geno_dir, "MZCER048.BAK2"))

# Copy new one over
file.copy(here("data/agriculture/MZCER048.CUL"),
          file.path(geno_dir, "MZCER048.CUL"))
```

##### Short season

```{r}
# Load scripts
source(here("scripts/read_sol_hor.R"))
source(here("scripts/run_dssat_perm.R"))

# Planting relevant
plant_density <- 3
row_spacing <- seq(75, 90, 5)
N_kg <- c(0, 10, 40, 80, 100)

# Short season
## Planting dates
planting_dates <- seq(19274, 19365, by = 15) # Oct to Dec

parameters <- expand.grid(
    start_year = 2019, 
    end_year = 2019, 
    planting_date = planting_dates, 
    plant_density = plant_density, 
    row_spacing = row_spacing, 
    N_kg = N_kg) %>% 
    mutate(id_param = 1:nrow(.))

cultivars <- cul$`@VAR#`

## Set a new class for the result       
yields <- do.call(rbind, lapply(1:nrow(samples), function(m) {
    x <- samples[m, ]
    do.call(rbind, lapply(1:nrow(parameters), function(n) {
        y <- parameters[n, ]
        planting_date <- y$planting_date
        soilid <- x$SoilProfil
        weatherid <- x$WeatherProfil
        coords <- data.frame(lat = x["lat"], lon = x["lon"])
        message(sprintf("(%s, %s), soil: %s, weather: %s.", 
                        x["lat"], x["lon"], soilid, weatherid))
        run_dssat_perm(
            soilid = soilid, plant_density = y$plant_density, 
            row_spacing = y$row_spacing, N_kg = y$N_kg,
            weatherid = weatherid, start_year = y$start_year, 
            end_year = y$end_year, planting_date = planting_date,
            coords = coords, cultivars = cultivars, 
            dssat_path = dssat_data_dir, 
            proj_path = here()) %>% mutate(
                id = x$id, adm1_pcode = x$adm1_pcode,
                soilid = soilid, plant_density = y$plant_density,
                row_spacing = y$row_spacing, N_kg = y$N_kg,
                weatherid = weatherid, planting_date = planting_date,
                lat = x[["lat"]], lon = x[["lon"]])
    }))
})) %>% mutate(cultivar = str_extract(EXNAME, "TZ[0-4]{4}")) %>% 
    select(id, adm1_pcode, lat, lon, soilid, weatherid, 
           cultivar, planting_date, plant_density, 
           row_spacing, N_kg, HWAH)

save(yields, 
     file = file.path(agro_result, "dssat_yield_ss_2019_20.rda"))

# Read survey data
nsca_yield <- read.csv(file.path(agro_dir, "nsca_district_yield.csv"))
nsca_yield <- right_join(nsca_yield, districts %>% st_drop_geometry(), 
                        by = c("region" = "Region"))

# Compare modeled and survey
yields <- left_join(yields, nsca_yield, by = "adm1_pcode")
save(yields, 
     file = file.path(agro_result, "dssat_survey_ss_2019_20.rda"))

best_yield <- do.call(rbind, lapply(unique(yields$region), function(rg){
    yield <- yields %>% filter(region == rg)
    real_yield <- unique(yield$yield)
    
    best_param <- yield %>% 
        group_by(cultivar, plant_density, row_spacing, 
                 N_kg, planting_date) %>% 
        summarise(yield = abs(mean(HWAH) / 1000 - real_yield)) %>% 
        ungroup() %>% filter(yield == min(yield))
    data.frame(region = unique(yield$region),
               yield = best_param$yield + real_yield)
})) %>% group_by(region) %>% 
    summarise(yield = mean(yield)) %>% ungroup()
```

##### Long season

```{r}
# Long season
## Planting dates
planting_dates <- seq(20122, 20152, by = 7) # March
```

